---
title: "Body Fat Tracker"
author: "Some Fat Guy"
date: "5/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(ggplot2)
  library(plotly)
  library(lubridate)
  library(gridExtra)
  library(functional)
  library(MASS)

  data <- fread("data.csv")
  data$bodyFatPct <- data$bodyFatPct / 100.0
  attach(data)

  targetFat <- 0.10
  firstWgt <- head(na.omit(weightLbs),1)
  lastWgt <- tail(na.omit(weightLbs),1)
  lastFat <- tail(na.omit(bodyFatPct),1)
  possibleWgts <- sort(c(160:195, lastWgt))

  BodyFatPct <- function(wgt) {
    fat <- lastWgt * lastFat
    muscle <- lastWgt - fat
    pmax(0, (wgt-muscle)/wgt)
  }

  HowMuchToLose <- function(goalFatPct, startWgt, startFatPct) {
    startFat = startWgt * startFatPct
    nonFatStart = startWgt - startFat
    goalWeight = nonFatStart/(1-goalFatPct)
    max(0, startWgt - goalWeight)
  }
  
  FinishLine <- function(goalFat, startWgt, startFat, deltaPerWeek=-1.5) {
    howMuch = HowMuchToLose(goalFat, startWgt, startFat)
    Sys.Date() - 7/deltaPerWeek * howMuch
  }

  pcts <- BodyFatPct(possibleWgts)
  dpcts <- c(NA, diff(pcts)/diff(possibleWgts))

  idx <- 1:nrow(data)
  w <- c(1000, rep(1,length(idx)-1)) #exp(-rev(idx-1)/21)
  s <- summary(lm(weightLbs ~ idx, data=data, weights=w))
  ratePerWeek <- coef(s)[2,1] * 7
  
  howMuch <- HowMuchToLose(targetFat, lastWgt, lastFat)
  doneDate <- FinishLine(targetFat, lastWgt, lastFat, ratePerWeek)

  pp <- function(x) sprintf("%0.1f%%", 100*x)
  pf <- function(x) sprintf("%0.1f", x)
```

## Fat Loss Summary
 - You are at `r pf(lastWgt) `lbs (`r pp(lastFat) ` fat).  
 - You have lost `r pf(firstWgt-lastWgt) `lbs over `r diff(ymd(range(date)))` days.
 - You are losing `r pf(-ratePerWeek) ` pounds per week.
 - You end weight target to hit `r pp(targetFat)` body fat needs to be `r pf(lastWgt-howMuch)`lbs. You have `r pf(howMuch)`lbs left to lose.
 - Estimated completion date is `r as.character(doneDate)`. (`r pf(as.numeric(doneDate - Sys.Date()))` days).

## Plots

```{r actFat, echo=FALSE}
  plt.wgt <- ggplot(mapping=aes(x=ymd(date))) + 
    geom_point(aes(y=weightLbs)) + 
    expand_limits(y=c(180, 195)) + 
    ggtitle("Weight over Time") + xlab("Date") + 
    stat_smooth(aes(y=weightLbs), method="loess")
                
  plt.actEstFat <- ggplot(mapping=aes(x=ymd(date))) + 
    geom_point(aes(y=BodyFatPct(weightLbs), color="estimated")) + 
    geom_point(aes(y=bodyFatPct, color="measured")) +
    geom_smooth(aes(y=BodyFatPct(weightLbs), color="estimated"), method="lm") + 
    xlab("Date") + ylab("Body Fat %") +
    ggtitle("Fat Pct Over Time") + scale_y_continuous(labels=pp)

  lab <- factor(ifelse(possibleWgts==lastWgt, "current", "path"))
  plt.fatWgt <- ggplot(mapping=aes(x=possibleWgts, y=pcts, color=lab)) + 
    xlab("Weight") + ylab("Body Fat %") +
    ggtitle("Estimated Fat Pct by Weight") + scale_y_continuous(labels=function(x) sprintf("%0.1f%%", 100*x)) + geom_point()
  
  plt.dfat <- ggplot(mapping=aes(x=possibleWgts, y=dpcts, color=lab)) + xlab("Weight") + ylab("Delta Body Fat %") +
    ggtitle("Body Fat % Change Per Pound Dropped") + scale_y_continuous(labels=function(x) sprintf("%0.2f%%", 100*x)) +
    geom_point()
  
  for (p in list(plt.wgt, plt.actEstFat, plt.fatWgt, plt.dfat))
    suppressWarnings(print(p))

```